---
title: "Sorting network data exploration"
output: 
  html_document: 
    keep_md: yes
    toc: true
    toc_float: true
    collapsed: false
    theme: default
  pdf_document: default
---

## Setup

### Dependencies

```{r}

library(readr)
library(tidyr)
library(ggplot2)
library(plyr)
library(reshape2)
library(cowplot)

source("https://gist.githubusercontent.com/benmarwick/2a1bb0133ff568cbe28d/raw/fb53bd97121f7f9ce947837ef1a4c65a73bffb3f/geom_flat_violin.R")
```

### Load Data

List out paths to relevant data files.
```{r}
min_networks_31250_SPT16_data_loc <- "../data/sn-exp-data/spt16/min_networks_31250.csv"
min_networks_1000K_SPT16_data_loc <- "../data/sn-exp-data/spt16/min_networks_1000000.csv"

min_networks_15625_SPT16_data_loc <- "../data/sn-exp-data/spt16/min_networks_15625.csv"
min_networks_500K_SPT16_data_loc <- "../data/sn-exp-data/spt16/min_networks_500000.csv"

min_networks_3125_SPT16_data_loc <- "../data/sn-exp-data/spt16/min_networks_3125.csv"
min_networks_100K_SPT16_data_loc <- "../data/sn-exp-data/spt16/min_networks_100000.csv"

min_networks_31250_SPT16_sol_summary_loc <- "../data/sn-exp-data/spt16/min_networks_31250__solutions_summary.csv"
min_networks_1000K_SPT16_sol_summary_loc <- "../data/sn-exp-data/spt16/min_networks_1000000__solutions_summary.csv"

min_networks_31250_SPT4_data_loc <- "../data/sn-exp-data/spt4/min_networks_31250.csv"
min_networks_1000K_SPT4_data_loc <- "../data/sn-exp-data/spt4/min_networks_1000000.csv"

min_networks_31250_SPT4_sol_summary_loc <- "../data/sn-exp-data/spt4/min_networks_31250__solutions_summary.csv"
min_networks_1000K_SPT4_sol_summary_loc <- "../data/sn-exp-data/spt4/min_networks_1000000__solutions_summary.csv"



```

Read in relevant data files.
```{r}
min_networks_31250 <- read.csv(min_networks_31250_SPT16_data_loc, na.strings = "NONE")
min_networks_1000K <- read.csv(min_networks_1000K_SPT16_data_loc, na.strings="NONE")

min_networks_15625 <- read.csv(min_networks_15625_SPT16_data_loc, na.strings = "NONE")
min_networks_500K <- read.csv(min_networks_500K_SPT16_data_loc, na.strings="NONE")

min_networks_3125 <- read.csv(min_networks_3125_SPT16_data_loc, na.strings = "NONE")
min_networks_100K <- read.csv(min_networks_100K_SPT16_data_loc, na.strings="NONE")

min_networks_31250_sol_summary <- read.csv(min_networks_31250_SPT16_sol_summary_loc, na.strings = "NONE")
min_networks_1000K_sol_summary <- read.csv(min_networks_1000K_SPT16_sol_summary_loc, na.strings="NONE")

# Load spt4 versions
min_networks_31250_SPT4 <- read.csv(min_networks_31250_SPT4_data_loc, na.strings = "NONE")
min_networks_1000K_SPT4 <- read.csv(min_networks_1000K_SPT4_data_loc, na.strings="NONE")

min_networks_31250_sol_summary_SPT4 <- read.csv(min_networks_31250_SPT4_sol_summary_loc, na.strings = "NONE")
min_networks_1000K_sol_summary_SPT4 <- read.csv(min_networks_1000K_SPT4_sol_summary_loc, na.strings="NONE")
```

## Analysis

### Generation 1,000,000

Once a solution is found, evolution is able to optimize. Less likely to find a solution in non-coevolution treatments here (cohorts).

```{r}
ggplot(data = filter(min_networks_1000K_sol_summary, uses_cohorts == "1"), mapping=aes(x=treatment, y=solutions_found, fill=treatment)) +
    geom_bar(stat="identity") + xlab("Condition") + ylab("Solutions Found") + guides(fill=FALSE) +
    geom_text(aes(label=solutions_found), nudge_y=1) +
    coord_flip() + ggtitle("Cohort lexicase selection - 1000K generations") + ggsave("cohorts_solution_cnt_1000K.png")

ggplot(data = filter(min_networks_1000K, uses_cohorts == "1"), mapping=aes(x=treatment, y=solution_size, fill=treatment, colour=treatment)) +
    geom_flat_violin(position=position_nudge(x=.2, y=0), adjust=2) +
    geom_point(position = position_jitter(width = .15), size = .25) +
    geom_boxplot(aes(x = treatment, y = solution_size),outlier.shape = NA, alpha = 0.3, width = .1, colour = "BLACK") +
    ylab("Solution Size") + ylim(60, 130) + xlab("Condition") + coord_flip() + guides(fill=FALSE, colour=FALSE) +
    ggtitle("Cohort lexicase selection - 1000K generations")

# + ggsave("cohorts_min_networks_1000K.png")
    
```

```{r}

ggplot(data = filter(min_networks_500K, uses_cohorts == "1"), mapping=aes(x=treatment, y=solution_size, fill=treatment, colour=treatment)) +
    geom_flat_violin(position=position_nudge(x=.2, y=0), adjust=2) +
    geom_point(position = position_jitter(width = .15), size = .25) +
    geom_boxplot(aes(x = treatment, y = solution_size),outlier.shape = NA, alpha = 0.3, width = .1, colour = "BLACK") +
    ylab("Solution Size") + ylim(60, 130) + xlab("Condition") + coord_flip() + guides(fill=FALSE, colour=FALSE) +
    ggtitle("Cohort lexicase selection - 500K generations")

ggplot(data = filter(min_networks_15625, uses_cohorts == "0"), mapping=aes(x=treatment, y=solution_size, fill=treatment, colour=treatment)) +
    geom_flat_violin(position=position_nudge(x=.2, y=0), adjust=2) +
    geom_point(position = position_jitter(width = .15), size = .25) +
    geom_boxplot(aes(x = treatment, y = solution_size),outlier.shape = NA, alpha = 0.3, width = .1, colour = "BLACK") +
    ylab("Solution Size") + ylim(60, 130) + xlab("Condition") + coord_flip() + guides(fill=FALSE, colour=FALSE) +
    ggtitle("full lexicase selection - 15625 generations")

# + ggsave("cohorts_min_networks_1000K.png")
    
```



```{r}

ggplot(data = filter(min_networks_100K, uses_cohorts == "1"), mapping=aes(x=treatment, y=solution_size, fill=treatment, colour=treatment)) +
    geom_flat_violin(position=position_nudge(x=.2, y=0), adjust=2) +
    geom_point(position = position_jitter(width = .15), size = .25) +
    geom_boxplot(aes(x = treatment, y = solution_size),outlier.shape = NA, alpha = 0.3, width = .1, colour = "BLACK") +
    ylab("Solution Size") + ylim(60, 130) + xlab("Condition") + coord_flip() + guides(fill=FALSE, colour=FALSE) +
    ggtitle("Cohort lexicase selection - 100K generations")

ggplot(data = filter(min_networks_3125, uses_cohorts == "0"), mapping=aes(x=treatment, y=solution_size, fill=treatment, colour=treatment)) +
    geom_flat_violin(position=position_nudge(x=.2, y=0), adjust=2) +
    geom_point(position = position_jitter(width = .15), size = .25) +
    geom_boxplot(aes(x = treatment, y = solution_size),outlier.shape = NA, alpha = 0.3, width = .1, colour = "BLACK") +
    ylab("Solution Size") + ylim(60, 130) + xlab("Condition") + coord_flip() + guides(fill=FALSE, colour=FALSE) +
    ggtitle("full lexicase selection - 3125 generations")

# + ggsave("cohorts_min_networks_1000K.png")
    
```


Use Fisher's exact

```{r}

# ctable <- min_networks_1000K %>% group_by(treatment, solution_found) %>% summarize(n=n()) %>% spread(solution_found, n)
# ctable[is.na(ctable)] <- 0
# m <- matrix(c(ctable$False, ctable$True), nrow=10, ncol=2, dimnames=list(ctable$treatment, c("True", "False")))
# fisher.test(m, alternative = "two.sided", workspace=1e6)
# 
# pairwise.fisher.test(min_networks_1000K_sol_summary$solutions_found, min_networks_1000K_sol_summary$total_runs, p.adjust.method = "bonferroni")
```

Use Kruskal-Wallis test to compare solution quality across cohort lexicase treatments (static tests, randomly changing tests, drifting tests, and coevolving tests).

```{r}
# df <- filter(min_networks_1000K, uses_cohorts == "1" & solution_found=="True")
# kruskal.test(solution_size ~ treatment, df)
# pairwise.wilcox.test(x=df$solution_size, g=df$treatment, paired=FALSE, p.adjust.method = "bonferroni")
```

### Generation 31,250

At generation 31,250 (where non-cohort treatments have had the same number of evaluations):

```{r}


# sol_cnts <- ggplot(data = min_networks_31250_sol_summary, mapping=aes(x=treatment, y=solutions_found, fill=treatment)) +
#           geom_bar(stat="identity") + xlab("Condition") + ylab("Solutions Found") + guides(fill=FALSE) +
#           geom_text(aes(label=solutions_found), nudge_y=1) +
#           coord_flip()
# sol_sizes <- ggplot(data = min_networks_31250, mapping=aes(x=treatment, y=solution_size, fill=treatment, colour=treatment)) +
#             geom_flat_violin(position=position_nudge(x=.2, y=0), adjust=2) +
#             geom_point(position = position_jitter(width = .15), size = .25) +
#             geom_boxplot(aes(x = treatment, y = solution_size),outlier.shape = NA, alpha = 0.3, width = .1, colour = "BLACK") +
#             + theme(axis.ticks.x=element_blank()) +
#             ylab("Solution Size") + ylim(60, 130) + xlab("Condition") + coord_flip() + guides(fill=FALSE, colour=FALSE)
# plot_grid(sol_sizes, sol_cnts)


ggplot(data = filter(min_networks_31250_sol_summary, uses_cohorts == "0"), mapping=aes(x=treatment, y=solutions_found, fill=treatment)) +
    geom_bar(stat="identity") + xlab("Condition") + ylab("Solutions Found") + guides(fill=FALSE) +
    geom_text(aes(label=solutions_found), nudge_y=1) +
    coord_flip() + ggtitle("Full Lexicase selection - 31,250 generations") + ggsave("full-lex_solution_cnt_31250.png")

ggplot(data = filter(min_networks_31250, uses_cohorts=="0"), mapping=aes(x=treatment, y=solution_size, fill=treatment, colour=treatment)) +
    geom_flat_violin(position=position_nudge(x=.2, y=0), adjust=2) +
    geom_point(position = position_jitter(width = .15), size = .25) +
    geom_boxplot(aes(x = treatment, y = solution_size),outlier.shape = NA, alpha = 0.3, width = .1, colour = "BLACK") +
    ylab("Solution Size") + ylim(60, 130) + xlab("Condition") + coord_flip() + guides(fill=FALSE, colour=FALSE) +
    ggtitle("Full lexicase selection - 31,250 generations") + ggsave("min_networks_31250_full-lex.png")
```

```{r}
# df <- filter(min_networks_31250, uses_cohorts == "0" & solution_found=="True")
# kruskal.test(solution_size ~ treatment, df)
# pwt<- pairwise.wilcox.test(x=df$solution_size, g=df$treatment, paired=FALSE, p.adjust.method = "bonferroni", exact=FALSE)
```



### Sorts per test = 4 (instead of 16)

Reduced number of test cases per test population.


```{r}
ggplot(data = filter(min_networks_1000K_sol_summary_SPT4, uses_cohorts == "1"), mapping=aes(x=treatment, y=solutions_found, fill=treatment)) +
    geom_bar(stat="identity") + xlab("Condition") + ylab("Solutions Found") + guides(fill=FALSE) +
    geom_text(aes(label=solutions_found), nudge_y=1) +
    coord_flip() + ggtitle("Cohort lexicase selection - 1000K generations \n SMALL TEST ORGS") + ggsave("cohorts_solution_cnt_1000K_small_tests.png")

ggplot(data = filter(min_networks_1000K_SPT4, uses_cohorts == "1"), mapping=aes(x=treatment, y=solution_size, fill=treatment, colour=treatment)) +
    geom_flat_violin(position=position_nudge(x=.2, y=0), adjust=2) +
    geom_point(position = position_jitter(width = .15), size = .25) +
    geom_boxplot(aes(x = treatment, y = solution_size),outlier.shape = NA, alpha = 0.3, width = .1, colour = "BLACK") +
    ylab("Solution Size") + ylim(60, 130) + xlab("Condition") + coord_flip() + guides(fill=FALSE, colour=FALSE) +
    ggtitle("Cohort lexicase selection - 1000K generations \n SMALL TEST ORGS") + ggsave("cohorts_min_networks_1000K_small_tests.png")
    
```

```{r}
ggplot(data = filter(min_networks_31250_sol_summary_SPT4, uses_cohorts == "0"), mapping=aes(x=treatment, y=solutions_found, fill=treatment)) +
    geom_bar(stat="identity") + xlab("Condition") + ylab("Solutions Found") + guides(fill=FALSE) +
    geom_text(aes(label=solutions_found), nudge_y=1) + 
    coord_flip() +  ggtitle("Full lexicase selection - 31250 generations \n SMALL TEST ORGS") + ggsave("full-lex_solution_cnt_1000K_small_tests.png")

ggplot(data = filter(min_networks_31250_SPT4, uses_cohorts=="0"), mapping=aes(x=treatment, y=solution_size, fill=treatment, colour=treatment)) +
    geom_flat_violin(position=position_nudge(x=.2, y=0), adjust=2) +
    geom_point(position = position_jitter(width = .15), size = .25) +
    geom_boxplot(aes(x = treatment, y = solution_size),outlier.shape = NA, alpha = 0.3, width = .1, colour = "BLACK") +
    ylab("Solution Size") + ylim(60, 130) + xlab("Condition") + coord_flip() + guides(fill=FALSE, colour=FALSE) +
    ggtitle("Full lexicase selection - 31250 generations \n SMALL TEST ORGS") + ggsave("min_networks_31250_full-lex_small_tests.png")
```